---
title: "Class Work"
author: "Nicholas Jacob"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(knitr)
library(ggplot2)
library(plotly)
library(doParallel)
registerDoParallel(8)
```

# Data
## Loading Data

```{r}
df <- read.csv("StudentSurveys.csv")

head(df)
```

## Wrangling the Data with `dplyr`

You should note that I added dplyr to the library above.  Following code gives me a frequency table.
```{r}
df |>
  count(Year)
```

We should note that there are `na`s in the data and the order is alphabetical.  We can change the datatype to fix this.

```{r}
df %>%
  mutate(Year = factor(Year,levels  = c("FirstYear","Sophomore", "Junior", "Senior")))%>%
  count(Year)
```

What if I want to look at GPA by grade level.

```{r}
df %>%
  group_by(Year)%>%
  summarise(avg = mean(GPA, na.rm = TRUE), std = sd(GPA, na.rm = TRUE))
```
```{r}
df |>
  filter(Year == "Junior")|>
 # mutate(HigherSAT = if_else(VerbalSAT >= MathSAT, "Verbal", "Math"))|> #if the actual wasn't there
  count(HigherSAT)

```

## Wrangling Data With Base

```{r}
table(df[df$Year=="Junior","HigherSAT"])
```

##Data Noise

```{r}
x = c(1:10)
y = x + rnorm(10)

plot(x,y, main = "Noisy Data")
```

```{r}
fit <- lm(y~x)

plot(x,y, main = "Linear Fit")
abline(fit)

```

```{r}
f <- approxfun(x, y)
curve(f(x), 0, 11, col = "green2")
points(x, y)

```

```{r lagrange interpolation}
lagrange <- function(x0, y0) {
  f <- function (x) {
    sum(y0 * sapply(seq_along(x0), \(j) {
      prod(x - x0[-j])/prod(x0[j] - x0[-j])
    }))
  }
  Vectorize(f, "x")
} #https://stackoverflow.com/questions/73994021/lagrange-interpolation-formula-in-r-programming

f <- lagrange(x,y)
plot(x,y)
curve(f, from = 1, to = 10, add = TRUE)
```
## Accuracy or Precision

How would you model?  I would have two random variables describing my desired outcome.  One I would think of as the initial placement and the other the noise.  I would tune both based on how high I allowed the accuracy and precision to be respectively.
```{r}
n <- 5
acc <- .1
prec <- .1

x<- rnorm(5,mean = 0, sd = acc)+rnorm(5,mean = 0, sd = prec)
y<- rnorm(5,mean = 0, sd = acc)+rnorm(5,mean = 0, sd = prec)

plot(x,y)#This doesn't work as well as I hoped...

```

## Data Types

```{r logical}
x <- c(1:100)

evens <- x%%2==0

x[evens]
```



Look at the student set, explore the different picks.
```{r}
disney <- read.csv("disney.csv")



weekdays(mdy(disney$Release.date..datetime.))[1]
```

## Statistics
```{r, echo = FALSE}
df |>
  group_by(Year) |>
  summarise("Aritmetic Mean" = mean(GPA, na.rm = TRUE), 
            "Geometric Mean" = exp(mean(log(GPA), na.rm = TRUE)),
            "Standard Deviation" = sd(GPA, na.rm = TRUE), 
            "InterQuartile Range" = IQR(GPA, na.rm = TRUE)
            )
```
Making this table look at bit nicer as it displays:

```{r, echo = TRUE}
df |>
  group_by(Year) |>
  summarise("Aritmetic Mean" = mean(GPA, na.rm = TRUE), 
            "Geometric Mean" = exp(mean(log(GPA), na.rm = TRUE)),
            "Standard Deviation" = sd(GPA, na.rm = TRUE), 
            "InterQuartile Range" = IQR(GPA, na.rm = TRUE)
            ) |>
  kable()

```



```{r}
t.test(df[df$Year == "FirstYear","GPA"],df[df$Year == "Senior","GPA"])
```


## Graphing 

I am going to completely ignore the base package way of graphing and stick to the `tidyverse` and `ggplot2`.

```{r, warning=FALSE}
ggplot(data = df, aes(x = Year, y = GPA)) +
  geom_boxplot(color = 'blue')
```
```{r}
ggplot(data = df, aes(x = Year))+
  geom_bar()
```

```{r}
ggplot(df, aes(x = Year, fill = Smoke)) +
  geom_bar()
```

```{r}



ggplot(df, aes(x = Year, y = Smoke, fill = GPA)) +
  geom_tile()
```

What if I really wanted those values?  It was much harder...
```{r}
df |> 
  group_by(Year, Smoke) |>
  summarise(AVG = mean(GPA, na.rm = TRUE)) |>
  ggplot(aes(x = Year, y = Smoke, fill = AVG)) +
  geom_tile() +
  geom_text(aes(label = round(AVG,2)))
```

```{r}
ggplot(df, aes(x = GPA))+
  geom_histogram() +
  labs(title = "Winner!")
```

```{r}
ggplot(df)+
  geom_qq(aes(sample = GPA)) +
  geom_qq_line(aes(sample = GPA))
```
```{r}
#, include = FALSE}
hist(df$TV, xlab = "TV",main = "Histogram of My TV Hours for Students")
```

```{r}
ggplot(df, aes(x = TV))+
  geom_histogram() +
  labs(title = "Histogram of TV time")
```

```{r}
ggplot(df, aes(x = GPA, y = Exercise))+
  geom_point()+
  geom_smooth(method = lm)
```

## Adding More to Your Graphics

Good graphs should have

1. Descriptive title
2. Appropriate Labels
3. Colors that are consistent and make sense
4. Allow the reader to glean the data quickly 
```{r}
ggplot(df, aes(x = GPA, y = Exercise,color = Gender))+
  geom_jitter()+
  geom_smooth(method = lm) +
  labs(title = "Is there Correlation of ", alt = "Graph revealing the correlation of female and males with gpa and exercise")+
  scale_color_manual(labels = c("Female","Male"), values = c('#FF5200','black' ))+
  theme_minimal()
```


```{r, message = FALSE, warning = FALSE, error = FALSE, echo = FALSE}
p <- df |>
  mutate(text = paste0("GPA: ", GPA, "\n", "Exercise: ", Exercise, "\n", "Gender: ",Gender, "\n", "What else?")) |>
  ggplot( aes(x = GPA, y = Exercise, text = text))+
  geom_jitter(aes(color = Gender))+
  geom_smooth(method = lm) +
  labs(title = "Is there Correlation of ", alt = "Graph revealing the correlation of female and males with gpa and exercise")+
  scale_color_manual(labels = c("Female","Male"), values = c('#FF5200','black' ))+
  theme_minimal() 

ggplotly(p,tooltip = "text")

```