---
title: "Class Work"
author: "Nicholas Jacob"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(knitr)
library(ggplot2)
library(plotly)
library(doParallel)
registerDoParallel(8)
library(boot)
library(caret)
library(ROCR)
library(MASS)
library(devtools)
```

# Data
## Loading Data

```{r}
df <- read.csv("StudentSurveys.csv")

head(df)
```

## Wrangling the Data with `dplyr`

You should note that I added dplyr to the library above.  Following code gives me a frequency table.
```{r}
df |>
  count(Year)
```

We should note that there are `na`s in the data and the order is alphabetical.  We can change the datatype to fix this.

```{r}
df %>%
  mutate(Year = factor(Year,levels  = c("FirstYear","Sophomore", "Junior", "Senior")))%>%
  count(Year)
```

What if I want to look at GPA by grade level.

```{r}
df %>%
  group_by(Year)%>%
  summarise(avg = mean(GPA, na.rm = TRUE), std = sd(GPA, na.rm = TRUE))
```
```{r}
df |>
  filter(Year == "Junior")|>
 # mutate(HigherSAT = if_else(VerbalSAT >= MathSAT, "Verbal", "Math"))|> #if the actual wasn't there
  count(HigherSAT)

```

## Wrangling Data With Base

```{r}
table(df[df$Year=="Junior","HigherSAT"])
```

##Data Noise

```{r}
x = c(1:10)
y = x + rnorm(10)

plot(x,y, main = "Noisy Data")
```

```{r}
fit <- lm(y~x)

plot(x,y, main = "Linear Fit")
abline(fit)

```

```{r}
f <- approxfun(x, y)
curve(f(x), 0, 11, col = "green2")
points(x, y)

```

```{r lagrange interpolation}
lagrange <- function(x0, y0) {
  f <- function (x) {
    sum(y0 * sapply(seq_along(x0), \(j) {
      prod(x - x0[-j])/prod(x0[j] - x0[-j])
    }))
  }
  Vectorize(f, "x")
} #https://stackoverflow.com/questions/73994021/lagrange-interpolation-formula-in-r-programming

f <- lagrange(x,y)
plot(x,y)
curve(f, from = 1, to = 10, add = TRUE)
```
## Accuracy or Precision

How would you model?  I would have two random variables describing my desired outcome.  One I would think of as the initial placement and the other the noise.  I would tune both based on how high I allowed the accuracy and precision to be respectively.
```{r}
n <- 5
acc <- .1
prec <- .1

x<- rnorm(5,mean = 0, sd = acc)+rnorm(5,mean = 0, sd = prec)
y<- rnorm(5,mean = 0, sd = acc)+rnorm(5,mean = 0, sd = prec)

plot(x,y)#This doesn't work as well as I hoped...

```

## Data Types

```{r logical}
x <- c(1:100)

evens <- x%%2==0

x[evens]
```



Look at the student set, explore the different picks.
```{r}
disney <- read.csv("disney.csv")



weekdays(mdy(disney$Release.date..datetime.))[1]
```

## Statistics
```{r, echo = FALSE}
df |>
  group_by(Year) |>
  summarise("Aritmetic Mean" = mean(GPA, na.rm = TRUE), 
            "Geometric Mean" = exp(mean(log(GPA), na.rm = TRUE)),
            "Standard Deviation" = sd(GPA, na.rm = TRUE), 
            "InterQuartile Range" = IQR(GPA, na.rm = TRUE)
            )
```
Making this table look at bit nicer as it displays:

```{r, echo = TRUE}
df |>
  group_by(Year) |>
  summarise("Aritmetic Mean" = mean(GPA, na.rm = TRUE), 
            "Geometric Mean" = exp(mean(log(GPA), na.rm = TRUE)),
            "Standard Deviation" = sd(GPA, na.rm = TRUE), 
            "InterQuartile Range" = IQR(GPA, na.rm = TRUE)
            ) |>
  kable()

```



```{r}
t.test(df[df$Year == "FirstYear","GPA"],df[df$Year == "Senior","GPA"])
```


## Graphing 

I am going to completely ignore the base package way of graphing and stick to the `tidyverse` and `ggplot2`.

```{r, warning=FALSE}
ggplot(data = df, aes(x = Year, y = GPA)) +
  geom_boxplot(color = 'blue')
```
```{r}
ggplot(data = df, aes(x = Year))+
  geom_bar()
```

```{r}
ggplot(df, aes(x = Year, fill = Smoke)) +
  geom_bar()
```

```{r}



ggplot(df, aes(x = Year, y = Smoke, fill = GPA)) +
  geom_tile()
```

What if I really wanted those values?  It was much harder...
```{r}
df |> 
  group_by(Year, Smoke) |>
  summarise(AVG = mean(GPA, na.rm = TRUE)) |>
  ggplot(aes(x = Year, y = Smoke, fill = AVG)) +
  geom_tile() +
  geom_text(aes(label = round(AVG,2)))
```

```{r}
ggplot(df, aes(x = GPA))+
  geom_histogram() +
  labs(title = "Winner!")
```

```{r}
ggplot(df)+
  geom_qq(aes(sample = GPA)) +
  geom_qq_line(aes(sample = GPA))
```
```{r}
#, include = FALSE}
hist(df$TV, xlab = "TV",main = "Histogram of My TV Hours for Students")
```

```{r}
ggplot(df, aes(x = TV))+
  geom_histogram() +
  labs(title = "Histogram of TV time")
```

```{r}
ggplot(df, aes(x = GPA, y = Exercise))+
  geom_point()+
  geom_smooth(method = lm)
```

## Adding More to Your Graphics

Good graphs should have

1. Descriptive title
2. Appropriate Labels
3. Colors that are consistent and make sense
4. Allow the reader to glean the data quickly 
```{r}
ggplot(df, aes(x = GPA, y = Exercise,color = Gender))+
  geom_jitter()+
  geom_smooth(method = lm) +
  labs(title = "Is there Correlation of ", alt = "Graph revealing the correlation of female and males with gpa and exercise")+
  scale_color_manual(labels = c("Female","Male"), values = c('#FF5200','black' ))+
  theme_minimal()
```


```{r, message = FALSE, warning = FALSE, error = FALSE, echo = FALSE}
p <- df |>
  mutate(text = paste0("GPA: ", GPA, "\n", "Exercise: ", Exercise, "\n", "Gender: ",Gender, "\n", "What else?")) |>
  ggplot( aes(x = GPA, y = Exercise, text = text))+
  geom_jitter(aes(color = Gender))+
  geom_smooth(method = lm) +
  labs(title = "Is there Correlation of ", alt = "Graph revealing the correlation of female and males with gpa and exercise")+
  scale_color_manual(labels = c("Female","Male"), values = c('#FF5200','black' ))+
  theme_minimal() 

ggplotly(p,tooltip = "text")

```

## Random Sampling

```{r}
i <- 0
n <- 1000
sampleSize <- 300

#sample(df$Exercise,sampleSize)
randomSampleMeans <- vector(,n) 
while(i<n){
  i<- i+1
  randomSampleMeans[i] = mean(sample(df$Exercise, sampleSize), na.rm = TRUE)
}

hist(randomSampleMeans)
```

```{r}
ttestobj <- t.test(df$Exercise, mu = 9)
ttestobj$stderr
```

```{r}
mean2 <- function(d,i){
  d <- d[i,]
  return(mean(d))
}
exercise <- df |>
  select(Exercise) |>
  filter(!is.na(Exercise)) 


bo <- boot(exercise,mean2,R = 1000)

plot(bo)
```

```{r}
dfNoNa <- df[complete.cases(df),]
fit <- trainControl(method = "repeatedcv",
                    number = 3,
                    repeats = 10)

train(Exercise ~ GPA,
      data = dfNoNa,
      method = "lm",
      trControl = fit)
```

# Linear Regression

## Correlation

```{r}
cor(dfNoNa$Height, dfNoNa$Weight)
```

Great, the black box works.  Let's check the value!

```{r}
xbar <- mean(dfNoNa$Height)
ybar <- mean(dfNoNa$Weight)

varx <- mean((dfNoNa$Height-xbar)^2)
vary <- mean((dfNoNa$Weight - ybar)^2)

cov <- mean((dfNoNa$Height-xbar)*(dfNoNa$Weight - ybar))

cov/(sqrt(varx*vary))
```

SWEET!

I should look at the scatterplot too.  I'll add the line that is the best as well.

```{r}
ggplot(dfNoNa, aes(x = Height, y = Weight, colour = Gender)) +
  geom_jitter()+
  geom_smooth(method = "lm")
```
```{r}
cor.test(dfNoNa$Height,dfNoNa$Weight)
```



A perfect fit
```{r}
x <- c(1:10)
y <- c(10:1)
fit <- lm(y~x)
plot(x,y)
abline(fit)
cor(x,y)
```

A noisy fit

```{r}
y <- -0.5*x + rnorm(10)
fit <- lm(y~x)
plot(x,y)
abline(fit)
cor(x,y)
```

Bad idea to fit

```{r}
y <- x^3

fit <- lm(y~x)
fit
plot(x,y)
abline(fit)
cor(x,y)
cor.test(x,y)
```

Got it!  Don't use the `poly` command.  It doesn't do what you'd expect...



## Regression

```{r}
lm(Weight ~ Height, data = dfNoNa)

```
Do it via formulas just to get it across...
```{r}
beta1 <- cov/varx
beta1

beta0 <- ybar - beta1*xbar
beta0
```

The lm, gives you more than just the coefficients.
```{r}
fit <- lm(Weight ~ Height, data = dfNoNa)

summary(fit)
```


Using `predict`.
```{r}
newData <- data.frame("Height" = c(69,42))
predict(fit, newData)
```

Making your own function
```{r}
line <- function(x){
   beta0+x*beta1
}
line(69)
```

The fit has lots of things to look at. 
```{r}
fit$coefficients["Height"]*69+fit$coefficients[1]
```

## Model Assessment
```{r}
plot(fit)
```

## Multiple Regression

```{r}
x <- c(1:10)
y <- x*x*x+3*x + rnorm(length(x))
x2 <- x^2
x3 <- x^3

fit <- lm(y~x + x2 +x3 +0-x2)
summary(fit)

fit$coefficients[2]
fun <- function(x){
  fit$coefficients[1]*x+fit$coefficients[2]*x*x*x
}

z <- apply(as.array(x),1,fun)
z
plot(x,z,"l")
points(x,y)
```


```{r}
fit2 <- lm(Weight ~ Height + Gender+Piercings+Award+BirthOrder +Smoke, data = dfNoNa)

summary(fit2)
```
## Categorical

One Hot Encode or Dummyfy


```{r}
dmy <- dummyVars(~.,df)
dfHot <- data.frame(predict(dmy, newdata = df))
head(dfHot)
```
Different regressions?

```{r}
lm(Weight ~ Height, data = dfNoNa[dfNoNa$Gender == "M",])
lm(Weight ~ Height, data = dfNoNa[dfNoNa$Gender != "M",])
```

Interactions can do that!

```{r}
lm(Weight ~ Height:Gender, dfNoNa)
lm(Weight ~ Height*Gender, dfNoNa)
```

Cyclical Encoding


```{r}
baby = data.frame(quarter = c(1,1,1,1,2,2,3,3,3,4,4,4,4,4,4,4,4,4))

n <- length(unique(baby$quarter))

tran <- function(x){
   c(cos(x*2*pi/n),sin(x*2*pi/n))
}


baby$x <- apply(as.array(baby$quarter),1,tran)[1,]
baby$y <- apply(as.array(baby$quarter),1,tran)[2,]

baby
```

Well it works...

## Extensions of Linear Model

```{r}
fitControl <- trainControl(method = "cv")
ridgeGrid <- expand.grid(lambda = seq(0.05,0.08,0.001))

ridgeCV <- train(Weight ~ ., data = dfNoNa,
      method = "ridge",
      trControl = fitControl,
      tuneGrid = ridgeGrid)

ridgeCV
plot(ridgeCV)
```

```{r}
ridgeCV$modelInfo
```


### Logistic Regression
```{r}
dfNoNa$Smoke <- factor(dfNoNa$Smoke)
fit <- glm(Smoke ~., data = dfNoNa, family = "binomial")

fit

sum(predict(fit, dfNoNa, type = 'response')>0.25)
```

```{r}
table(dfNoNa$Smoke)
```

## Principle Components

```{r}
dfComplete <- dfHot[complete.cases(dfHot),]

pca <- prcomp(dfComplete)

plot(pca)

summary(pca)
```

```{r}
fitControl <- trainControl(method = "cv")
ncompGrid <- expand.grid(ncomp = c(1:20))

train(Weight ~., data = dfNoNa,
      method = "pcr",
      tuneGrid = ncompGrid)
```
# Classification

```{r}
dfNoNa$Gender <- as.factor(dfNoNa$Gender)

fitControl <- trainControl(method = "LGOCV", p = 0.66, number = 1)
classtrain <- train(Gender~Height + Weight, data = dfNoNa,
      method = "knn",
      trControl = fitControl,
      tuneGrid = expand.grid(k=1:9))

classtrain$
```
The best one was `{r} classtrain$bestTune$k`  

```{r}
confusionMatrix(classtrain)
```
```{r}
a <- createDataPartition(dfNoNa$Gender, list=FALSE, p = 0.66)
train <- dfNoNa[a,]
test <- dfNoNa[-a,]

knnfit <- knn3(Gender ~ Weight +Height, data = train)

pred <- predict(knnfit,test, type = "class")

confusionMatrix(table(predicted =pred,actual = test$Gender ))
```
```{r}
pred <- predict(knnfit,test)
pred
1*(test$Gender == "F")
data.frame(pred)$F
data.frame(pred)$F
pred <- prediction(data.frame(pred)$F, 1*(test$Gender == "F"))
pref <- performance(pred, "tpr", "fpr")
plot(pref, colorize = T)
```

```{r}
data(ROCR.simple)
df <- data.frame(ROCR.simple)
pred <- prediction(df$predictions, df$labels)
perf <- performance(pred,"tpr","fpr")
pred
plot(perf,colorize=TRUE)
```
## Decision Tree

```{r}
library(rpart)
library(rpart.plot)

tree <- rpart(Gender ~ Height + Weight,data = dfNoNa)
tree
```
```{r tree viz}
rpart.plot(tree, extra = 2)
```
```{r}
pred <- predict(tree, dfNoNa, type = "class")
head(pred)
```
Each has been classified into its category.  You can also recover the probabilities of the classification by dropping the *type = "class"*
```{r}
predict(tree, dfNoNa) %>%
  head()
```


Confusion table follows by using the classified data.

```{r}
confusion_table <- with(dfNoNa, table(Gender, pred))
confusion_table
```

or

```{r}
confusionMatrix(table(predicted =pred,actual = dfNoNa$Gender ))
```

```{r}
varImp(tree)
```

```{r}
tree <- rpart(Gender ~ . ,data = dfNoNa, method = "class",
              control = rpart.control(minsplit = 2, cp = 0.0001)
              )
rpart.plot(tree, extra = 2)
```

```{r}
varImp(tree)
```

```{r}
fitControl <- trainControl(method = "LGOCV", p = 0.66, number = 1)
rpartGrid <- expand.grid(cp = seq(0,1,0.1))
classtrain <- train(Gender~., data = dfNoNa,
      method = "rpart",
     # trControl = fitControl,
      tuneGrid = rpartGrid)

classtrain
```

## k Nearest Neighbors

```{r}
shrinkdata <- createDataPartition(dfNoNa$Year , p = 0.1, list = FALSE)
model <- knn3(Gender ~ Height + Weight, data = dfNoNa[shrinkdata,], k = 5)
pred <- predict(model, dfNoNa, type = "class")


plot(dfNoNa[shrinkdata,"Height"],
     dfNoNa[shrinkdata,"Weight"],
     col = dfNoNa[shrinkdata,"Gender"], 
     pch =1*(pred[shrinkdata]=="F") )

```
PCA does much better but is really damn hard to code...

```{r}
head(dfNoNa)
dv <- dummyVars(~., data = dfNoNa[,!(colnames(dfNoNa) %in% c("Gender"))])
dfNoNaDummy <- predict(dv, dfNoNa)
pca <- princomp(dfNoNaDummy, cor =TRUE)
plot(pca)
dfPCA <- data.frame(predict(pca,dfNoNaDummy))
dfPCA$Gender = dfNoNa$Gender
model <- knn3(Gender ~., data = dfPCA)
pred <- predict(model, dfPCA, type = "class")
confusionMatrix(table(predicted =pred,actual = dfNoNa$Gender ))
```
## Logistic Regression
```{r}
dfNoNa$GenderBool <- 1*(dfNoNa$Gender == "M")

sillyModel <- lm(GenderBool~Weight, data = dfNoNa)

plot(dfNoNa$Weight, dfNoNa$GenderBool)
abline(sillyModel)
```

Now I'll actually do logistic regression!
```{r}
logit <- glm(Gender ~ Weight, family = binomial, data = dfNoNa)
p <- predict(logit, data = dfNoNa, type = "response")
pr <- prediction(p, dfNoNa$Gender)
plot(performance(pr, measure = "tpr", x.measure = "fpr"))
```

Above I plotted the ROC curve.  I'd like to recreate the graphic above and show how the probabilities change.

```{r}
fakedf <- data.frame(Weight = seq(min(dfNoNa$Weight),max(dfNoNa$Weight), 0.5))
fake <- predict(logit,fakedf, type = "response")

plot(dfNoNa$Weight, dfNoNa$GenderBool)
lines(fakedf$Weight,fake)

```

### LDA
```{r}
linear<- lda(Award ~Weight+Height+Gender, dfNoNa)
linear
```
```{r}
p <- predict(linear, dfNoNa)
ldahist(p$x[,1], g = dfNoNa$Award)
```



```{r}
p2 <- predict(linear,dfNoNa)$class
confusionMatrix(table(Predicts = p2, Actual = dfNoNa$Gender))
```

### Support Vector Machines

```{r}
svmGrid <- expand.grid(C = seq(0.04,0.1,0.005))

train(Gender ~ Height +Weight, data = dfNoNa,
      method = "svmLinear",
      trControl = trainControl(method = "cv", p = 0.33),
      tuneGrid = svmGrid)
```

## Clustering
```{r}
cl <- kmeans(dfNoNa[,c("Height","Weight")],3)


plot(dfNoNa[,c("Height","Weight")], col = cl$cluster)# pch =dfNoNa$Gender)
points(cl$centers, col = 1:3, pch = 8, cex = 2)
```

Was 3 the correct number of clusters?
```{r}
library(factoextra)
fviz_nbclust(x = dfNoNa[,c("Height","Weight")],FUNcluster = kmeans, method = 'wss' )
```

```{r}

kmeans_res = kmeans(dfNoNa[,c("Height","Weight")], 2)
fviz_cluster(kmeans_res, dfNoNa[,c("Height","Weight")])
# determining k using the silhouette method
fviz_nbclust(x = dfNoNa[,c("Height","Weight")],FUNcluster = kmeans, method = 'silhouette' )
```

```{r, warning=FALSE}
#not sure what this does...
set.seed(42)
library(cluster)
clust_gap <- clusGap(x = dfNoNa[,c("Height","Weight")], FUN = kmeans, K.max = 15, nstart = 25,  B = 5)
fviz_gap_stat(clust_gap)
```
```{r}
ag1 <- agnes(dfNoNa[,c("Height","Weight")])
plot(ag1)
```

### DBSCAN

```{r}
library(dbscan)

res <- dbscan(dfNoNa[,c("Height", "Weight")], eps = 2)
plot(dfNoNa[,c("Height", "Weight")], col = res$cluster)
points(dfNoNa[res$cluster == 0,c("Height", "Weight")], pch = 3, col = "grey")
```

```{r}
clplot(dfNoNa[,c("Height", "Weight")],res)
```

```{r}
hullplot(dfNoNa[,c("Height", "Weight")],res)
```

```{r}
x <- scale(dfNoNa[,c("Height", "Weight")])

res <- dbscan(x, eps = 0.4)
clplot(x,res)
```