---
title: "Class Work"
author: "Nicholas Jacob"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
```

# Data
## Loading Data

```{r}
df <- read.csv("https://raw.githubusercontent.com/nurfnick/Data_Mining_R/refs/heads/main/StudentSurveys.csv")

head(df)
```

## Wrangling the Data with `dplyr`

You should note that I added dplyr to the library above.  Following code gives me a frequency table.
```{r}
df |>
  count(Year)
```

We should note that there are `na`s in the data and the order is alphabetical.  We can change the datatype to fix this.

```{r}
df %>%
  mutate(Year = factor(Year,levels  = c("FirstYear","Sophomore", "Junior", "Senior")))%>%
  count(Year)
```

What if I want to look at GPA by grade level.

```{r}
df %>%
  group_by(Year)%>%
  summarise(avg = mean(GPA, na.rm = TRUE), std = sd(GPA, na.rm = TRUE))
```
```{r}
df |>
  filter(Year == "Junior")|>
 # mutate(HigherSAT = if_else(VerbalSAT >= MathSAT, "Verbal", "Math"))|> #if the actual wasn't there
  count(HigherSAT)

```

## Wrangling Data With Base

```{r}
table(df[df$Year=="Junior","HigherSAT"])
```

##Data Noise

```{r}
x = c(1:10)
y = x + rnorm(10)

plot(x,y, main = "Noisy Data")
```

```{r}
fit <- lm(y~x)

plot(x,y, main = "Linear Fit")
abline(fit)

```

```{r}
f <- approxfun(x, y)
curve(f(x), 0, 11, col = "green2")
points(x, y)

```

```{r lagrange interpolation}
lagrange <- function(x0, y0) {
  f <- function (x) {
    sum(y0 * sapply(seq_along(x0), \(j) {
      prod(x - x0[-j])/prod(x0[j] - x0[-j])
    }))
  }
  Vectorize(f, "x")
} #https://stackoverflow.com/questions/73994021/lagrange-interpolation-formula-in-r-programming

f <- lagrange(x,y)
plot(x,y)
curve(f, from = 1, to = 10, add = TRUE)
```
## Accuracy or Precision

How would you model?  I would have two random variables describing my desired outcome.  One I would think of as the initial placement and the other the noise.  I would tune both based on how high I allowed the accuracy and precision to be respectively.
```{r}
n <- 5
acc <- .1
prec <- .1

x<- rnorm(5,mean = 0, sd = acc)+rnorm(5,mean = 0, sd = prec)
y<- rnorm(5,mean = 0, sd = acc)+rnorm(5,mean = 0, sd = prec)

plot(x,y)#This doesn't work as well as I hoped...

```

## Data Types

```{r logical}
x <- c(1:100)

evens <- x%%2==0

x[evens]
```



Look at the student set, explore the different picks.
```{r}
disney <- read.csv("disney.csv")



weekdays(mdy(disney$Release.date..datetime.))[1]
```

## Statistics
```{r}
df |>
  group_by(Year) |>
  summarise("Aritmetic Mean" = mean(GPA, na.rm = TRUE), 
            "Geometric Mean" = exp(mean(log(GPA), na.rm = TRUE)),
            "Standard Deviation" = sd(GPA, na.rm = TRUE), 
            "InterQuartile Range" = IQR(GPA, na.rm = TRUE)
            )
```

