---
title: "Easy Data Loading"
author: "Nicholas Jacob"
date: "8/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## From Github

The dataset I am going to look at includes all people killed by a police officer in 2016.  I gathered this dataset a while back and put it on my github.  You can see the original from the Gaurdian Newspaper [here](https://www.theguardian.com/us-news/ng-interactive/2015/jun/01/the-counted-police-killings-us-database).  I uploaded the file to my github and created a `raw` link for it.  I'll now read it into my code

```{r load the data from link}
df <- read.csv("https://raw.githubusercontent.com/nurfnick/Data_Sets_For_Stats/master/CuratedDataSets/the-counted-2016.csv")
```
I give a small sample of the data using the head command.  Printing all of the data should be avoided!

```{r}
head(df)
```

## From Upload to RStudio

If you were not going to load the data from a link (github only supports small data files).  You can load directly into RStudio.  Use `Upload` and select our file, I did this with a file called **disney.xlsx**.  Now that it is in, I'll load it into R.

```{r Load Excel From Directory}
library(readxl)
df2 = read_excel("disney.xlsx")
```

```{r}
head(df2)
```
## Built Into R

There are some datasets included in packages for R.  You can find a list [here](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/00Index.html)  I'll load something from the list

```{r}
Titanic
```
This will *NOT* work for the project...`iris` or `cars` might...

## A Smattering of R Packages 

I found mention of a baseball stats package so I've added to my library by installing.  You can find more info [here](http://www.seanlahman.com/baseball-archive/statistics/)

```{r}
library(Lahman)
#help(Lahman)
head(Batting)
```

```{r}
library(babynames)
head(babynames)
```
```{r}
library(nycflights13)
head(flights)
```
```{r}
library(fec16)
head(candidates)
```