---
title: "Earthquakes 2021"
author: "Nicholas Jacob"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(tidyverse)

url = 'https://en.wikipedia.org/wiki/List_of_earthquakes_in_2021'
```

## Scrapping Wikipedia

First I am going to gather the data from wikipedia with the **rvest** library.

```{r gatherhtml}
html <- read_html(url)
```
Let's take a look at the html and see what we have gathered.

```{r veiwhtml}
html
```
Look's like we got something.  Let's see if we can extract the tables from the webpage.

```{r tableshtml}
html %>%
  html_table()
```
That was a lot of tables!  Let's dig in and see if we can extract the monthly ones.  I think the seventh one is January's.
```{r}
tables <- html %>%
  html_table(header = TRUE)

tables[[7]][-1,]
```
Let's now create our eventual final dataframe and make certain we add a column highlighting what month this happened in.

```{r}
jan <- tables[[7]][-1,]

df <- jan %>% add_column(Month = 'January')


```
Let's automate the rest.  It is currently August so there is only 8 months of data.

```{r}
monthlist = c('January','February','March','April','May','June','July','August')
counter = c(2:8)

for (number in counter){
  newtable <- tables[[5+2*number]][-1,] #the minus one is to get rid of the title of the columns 
  newtable <- newtable %>% add_column(Month = monthlist[number])
  df <- bind_rows(df,newtable)
}
```
```{r}
head(df)
```
I am going to need to do some cleaning now that I have gathered the data.  In the *Date* column there is a reference hyperlink in square brackets.  Let's get rid of that and rename it *Day*
```{r cleanDay}
df <- df %>%
  mutate(
    Day = str_remove_all(Date,pattern = "\\[*[0-9]*\\]")
  )
  
```

Next I'd like to change all the casualty entries from *NA* to zero.

```{r cleanCasualties}
df <- df %>%
  mutate(
    Deaths = str_replace_all(Casualties...7,"\\-","0"),
    Injuries = str_replace_all(Casualties...8,"\\-","0")
  )
```

They were not actually *NA* but were -.  Tricky!

Let's continue looking to extract some information about country and location
```{r}
df<- df %>%
  mutate(
    Offshore = str_detect(`Country and location`,"offshore"),
    Country = str_extract(`Country and location`,'[A-z]+')
  )
```


A couple of these are wrong, I'll fix the obvious ones and come back if I see others I missed later.

```{r}
df <- df %>%
  mutate(
    Country = replace(Country, Country == 'United','United States'),
    Country = replace(Country, Country == 'South', 'South Georgia'),
    Country = replace(Country, Country == 'Papua','Papua New Guinea')
  )

```


```{r}
df <- df %>%
  mutate(
    Country = replace(Country, str_detect(`Country and location`,'New Zealand'),'New Zealand'),
    Country = replace(Country, str_detect(`Country and location`,'New Cal'),'New Caledonia')
                 )
```

```{r}
df <- df %>%
  rename(Depth = 'Depth (km)')
```
Okay I think I have extracted all the info I am going to.  I'll clean up the dataset and oranize it.
```{r organize}
df <- df %>%
  select(c(Month,Day,Country,Deaths,Injuries,Mw,Depth,MMI,Offshore))
```
```{r}
df <- df %>%
  mutate(
    Deaths = str_remove_all(Deaths,","),
    Injuries = str_remove_all(Injuries, ",")
  )
```

```{r}
df <- df %>%
  mutate(Deaths = as.integer(Deaths),
         Injuries = as.integer(Injuries),
         Depth = as.numeric(Depth),
         Day = as.integer(Day),
         Mw = as.numeric(Mw))

```

## Ready For EDA

```{r}
df %>%
  summarize(mean())
```
