---
title: "Association on Earthquakes"
author: "Nicholas Jacob"
date: "9/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(tidyverse)

url = 'https://en.wikipedia.org/wiki/List_of_earthquakes_in_2021'

html <- read_html(url)

tables <- html %>%
  html_table(header = TRUE)


jan <- tables[[7]][-1,]

df <- jan %>% add_column(Month = 'January')



monthlist = c('January','February','March','April','May','June','July','August','September','October')
counter = c(2:10)

for (number in counter){
  newtable <- tables[[5+2*number]][-1,] #the minus one is to get rid of the title of the columns 
  newtable <- newtable %>% add_column(Month = monthlist[number])
  df <- bind_rows(df,newtable)
}

df <- df %>%
  mutate(
    Day = str_remove_all(Date,pattern = "\\[*[0-9]*\\]")
  )
  

df <- df %>%
  mutate(
    Deaths = str_replace_all(Casualties...7,"\\-","0"),
    Injuries = str_replace_all(Casualties...8,"\\-","0")
  )

df<- df %>%
  mutate(
    Offshore = str_detect(`Country and location`,"offshore"),
    Country = str_extract(`Country and location`,'[A-z]+')
  )

df <- df %>%
  mutate(
    Country = replace(Country, Country == 'United','United States'),
    Country = replace(Country, Country == 'South', 'South Georgia'),
    Country = replace(Country, Country == 'Papua','Papua New Guinea')
  )


df <- df %>%
  mutate(
    Country = replace(Country, str_detect(`Country and location`,'New Zealand'),'New Zealand'),
    Country = replace(Country, str_detect(`Country and location`,'New Cal'),'New Caledonia')
                 )

df <- df %>%
  rename(Depth = 'Depth (km)')

df <- df %>%
  select(c(Month,Day,Country,Deaths,Injuries,Mw,Depth,MMI,Offshore))

df <- df %>%
  mutate(
    Deaths = str_remove_all(Deaths,","),
    Injuries = str_remove_all(Injuries, ",")
  )

df <- df %>%
  mutate(Deaths = as.integer(Deaths),
         Injuries = as.integer(Injuries),
         Depth = as.numeric(Depth),
         Day = as.integer(Day),
         Mw = as.numeric(Mw))


df <- df %>%
  drop_na()

df <- df %>% mutate(
  Offshore = factor(Offshore == TRUE, levels = c(TRUE, FALSE),
                    labels = c('offshore','on land' )),
  MMI = factor(MMI, levels = c("-","I","II","III","IV","V","VI","VII","VIII","IX"), labels = c(0,"I","II","III","IV","V","VI","VII","VIII","IX")) ,
  Month = factor(Month),
  Country = factor(Country)
)
```



I will use the same data from wikipedia on earthquakes in 2021 that I gathered [here](https://en.wikipedia.org/wiki/List_of_earthquakes_in_2021).  The previous portion of this project was published [here](https://nurfnick.github.io/Data_Mining_R/EarthQuakes.html)  I have not included all the data cleaning in this document even though it was all necessary.  It is hidden in the first cell!
```{r}

tail(df)
```

## Association

```{r new libraries}

library(arules)
library(arulesViz)
```

Added some new libraries.  And now to convert the data for association analysis.

```{r}
transactions(df)
```

Lot's of errors here!  Let's look at the columns and see if we can convert them into factors (or Boolean) for analysis.
```{r}
colnames(df)[c(2,4,5,6,7)]
```

I don't think I can do anything with *Day* *Mw* or *Depth* but I could ask if there were any *Deaths* or *Injuries* since most were zeroed out in my data.

```{r}
df <- df %>% mutate(
  Deaths = (Deaths > 0),
  Injuries = (Injuries >0)
)
```

Then I'll run the transactions again and see how it has cleaned up.

```{r}
trans <- transactions(df)
```
I am happy to let R do the default discretization of the rest of the data.  Partly because what is left I could not come up with better cutoffs.  Depth, Mw and Day nothing strikes me as important.  Day might be important in stock data but for a natural phenomenon, I can see no reason to chop it differently.


```{r}
summary(trans)
```

```{r}
colnames(trans)
```

```{r}
inspect(trans[1:3])
```

```{r}
image(trans)
```

```{r}
itemFrequencyPlot(trans,topN = 20)
```
```{r}
vertical <- as(trans, "tidLists")
as(vertical, "matrix")[1:10, 1:5]
```

```{r}
trans
trans_Nepal <- trans[trans %in% "Country=Nepal"]
trans_Nepal
```

```{r}

inspect(trans_Nepal)
```
Looks like there was only one earthquake in Nepal during May and it did cause injuries.

Let's find some frequent itemsets

```{r}
its <- apriori(trans, parameter=list(target = "frequent"))
```

```{r}
its
```

```{r}
inspect(head(its, n = 10))
```

```{r}
ggplot(tibble(`Itemset Size` = factor(size(its))), aes(`Itemset Size`)) + geom_bar()
```

Now we'll build some rules.  Very similar to looking at frequent!
```{r}
rules <- apriori(trans, parameter = list(support = 0.05, confidence = 0.9))
inspect(head(rules))
```

```{r}
plot(rules,jitter = 1)
```

```{r}
plot(rules, shading = "order")
```