---
title: "Earthquakes 2021"
author: "Nicholas Jacob"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(tidyverse)
url = 'https://en.wikipedia.org/wiki/List_of_earthquakes_in_2021'
```

## Scrapping Wikipedia

First I am going to gather the data from wikipedia with the **rvest** library.

```{r gatherhtml}
html <- read_html(url)
```
Let's take a look at the html and see what we have gathered.

```{r veiwhtml}
html
```
Look's like we got something.  Let's see if we can extract the tables from the webpage.

```{r tableshtml}
html %>%
  html_table()
```
That was a lot of tables!  Let's dig in and see if we can extract the monthly ones.  I think the seventh one is January's.
```{r}
tables <- html %>%
  html_table(header = TRUE)

tables[[7]][-1,]
```
Let's now create our eventual final dataframe and make certain we add a column highlighting what month this happened in.

```{r}
jan <- tables[[7]][-1,]

df <- jan %>% add_column(Month = 'January')


```
Let's automate the rest.  It is currently August so there is only 8 months of data.

```{r}
monthlist = c('January','February','March','April','May','June','July','August')
counter = c(2:8)

for (number in counter){
  newtable <- tables[[5+2*number]][-1,] #the minus one is to get rid of the title of the columns 
  newtable <- newtable %>% add_column(Month = monthlist[number])
  df <- bind_rows(df,newtable)
}
```
```{r}
df
```
I am going to need to do some cleaning now that I have gathered the data.  In the *Date* column there is a reference hyperlink in square brackets.  Let's get rid of that and rename it *Day*
```{r cleanDay}
df <- df %>%
  mutate(
    Day = str_remove_all(Date,pattern = "\\[*[0-9]*\\]")
  )
  
```

Next I'd like to change all the casualty entries from *NA* to zero.

```{r cleanCasualties}
df <- df %>%
  mutate(
    Deaths = str_replace_all(Casualties...7,"\\-","0"),
    Injuries = str_replace_all(Casualties...8,"\\-","0")
  )
```